### 2023-04-17 definitions.tex
details

% We prefer to use infix notation and $(s,a ,s')\in {\rightarrow} $ will be abbreviated as    $s  \xrightarrow{a}  s'$ to represent a transition from configuration $s$ to configuration $s'$ with label $a$. 
We prefer to use infix notation and write $s \rightarrow s'$ to denote a {\em transition} from state $s$ to state $s'$ (i.e., $ (s,s') \in  {\rightarrow} $).
We write $\rightarrow^{k}$, $\rightarrow^{+}$, $\rightarrow^{=}$, $\rightarrow^{*}$
for the $k$-step iteration of $\rightarrow$, its transitive closure, its reflexive closure, its reflexive and transitive closure.
Let $X,Y \subseteq S$ and $k \in \mathbb{N}$. We denote $X \longrightarrow^{*} Y$ (resp. $X \longrightarrow^{\leq k} Y$) if from all states $x \in X$ there exists a path (resp. of length smaller than $k$) that reaches a state $y \in Y$.
\noindent
The set of {\em (immediate) successors} of a state $s \in S$ is defined as 
 $\post(s) = \{ s' \in S \mid  ~ s \xrightarrow{} s'\}$. 
% $\post(s) = \{ s' \in S \mid  ~ \exists a \in \Sigma~ s \xrightarrow{a} s'\}$. 
A state without successors is called a {\em dead end}. 
The set of {\em (immediate) predecessors} of a state $s \in S$ is deﬁned as
 $\pred(s) = \{ s' \in S \mid  ~ s' \xrightarrow{} s\}$. 
% $\pred(s) = \{ s' \in S \mid  ~ \exists a \in \Sigma~ s' \xrightarrow{a} s\}$. 
By iterating $\pred$ and $\post$ we obtain  
$\post^n(s) = \{ s' \in S \mid  ~ s' \xrightarrow{}^n s\}$
and
$\pred^n(s) = \{ s' \in S \mid  ~ s' \xrightarrow{}^n s\}$.
However, we are generally more interested in
$\post^{\leq n}(s) = \bigcup_{1 \leq i \leq n} \post^i(s)$
and
$\pred^{\leq n}(s) = \bigcup_{1 \leq i \leq n} \pred^i(s)$. \\

@@

\begin{definition}\cite{DBLP:journals/iandc/Finkel90,DBLP:journals/tcs/FinkelS01}
A {\em well-structured transition system} (WSTS)  $\mathscr{S}=(S, \rightarrow, \leq)$
is a transition system $(S, \rightarrow)$
equipped with a wqo ${\leq} \subseteq S \times S$ such that  
%  decidable ≤  wqo on S, i.e., for each two given states s, s ′ ∈ S, it is decidable whether s ≤ s ′ . 
the transition relation $ \rightarrow$ is compatible with $\leq$, i.e., for all 
$s_1, t_1 , s_2 \in S$
	with $s_1 \leq s_2$  and $s_1 \rightarrow t_1$, there exists 
	$t_2 \in S$ with 
	$t_1 \leq t_2$ and $s_2 \rightarrow^{*} t_2$.
\end{definition}

@@

The following property is very usefull: if $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$ is upward-closed, then $\pred(U )$, $\pred^k(U )$ with $k\geq0$, and $\pred^*(U )$ are upward-closed \cite{DBLP:journals/tcs/FinkelS01}.
%\alain{étrange ces deux propositions, l'une suffirait}
\iffalse
Proof. Assume $s \in \pred (I )$. Then $s \rightarrow t$ for some $t \in I $. If now $s' \geq s$ then strong upward-compatibility entails that $s' \rightarrow t'$ for some $t' \geq t$. Then $t' \in I$ and $s' \in \pred(I )$.

\mathieu{On peut probablement enlever la Proposition $8$ alors. La proposition $9$ étant plus directement pertinente pour les lemmes/preuves (permet d'avoir itérativement $\pred^k(I)$ upward-closed if $I$ upward-closed - for all $k$).}
\fi
%
Now, we are interested to know a simple condition that insures that a finite basis of $\pred^*(\uparrow s )$ is computable for every $s \in S$.

\begin{definition}\cite{DBLP:journals/tcs/FinkelS01,DBLP:journals/iandc/AbdullaCJT00}
A WSTS $\mathscr{S}$ has {\em effective pred-basis} if there exists an algorithm accepting
any state $s \in S$ and returning $pb(s)$, a finite basis of $\uparrow \pred(\uparrow s)$.
\end{definition}

\begin{theorem}(ref)
A finite basis of $ \pred^*(U)$ is computable for any effective WSTS $\mathscr{S}$ with effective pred-basis and any upward-closed set $U \subseteq S$ given with its finite basis $B_U$.
\end{theorem}

@@

% (ii) Given a finite set $A \subseteq S$ with $I =\uparrow A$, we can compute a finite basis $B$ of $I$.
% \alain{(ii) est bizarre car A est déjà une base finie de $\uparrow A$, et $\min(A)$ est la base minimale}
%			\end{fact}

For a WSTS $\mathscr{S}$ and an upward-closed set $U$, let us study the convergence of the sequence defined by $U_0=U$ and $U_k= U \cup \pred(U_{k-1})$ for $k \geq 1$. When $\mathscr{S}$ has the strong compatibility, the sets $U_k$ are upward closed and $U_k \subseteq U_{k+1}$ so we know that the sequence $(U_k)_k$ converges. Let us define the \emph{index} of convergence of the sequence $U_k$ as the smallest $k_0$ s.t. $U_k = U_{k_0}$ for all $k \geq k_0$. We then have:  $\pred^*(U) = U_{k_0}$. \\
%

\noindent
\subsection{Computing the maximal $X$-live system}

Let $\mathscr{S}=(S, \rightarrow, \leq,s_0)$ be an (ordered transition) system with initial state $s_0$ and $X \subseteq S$ be an upward-closed set given by its finite minimal basis $B_X$. We say that $\mathscr{S}=(S, \rightarrow, \leq,s_0)$ is $X$-live if from any reachable state $s$ (from $s_0$), there exists (at least) a run that goes into $X$: $s \xrightarrow{*} X$.  We may verify that  $\mathscr{S}=(S, \rightarrow, \leq,s_0)$ is $X$-live iff  $\mathscr{S}=(S, \rightarrow, \leq,s_0)$ is $\pred^*(X)$-live. The \emph{maximal $X$-live subsystem} of $(S,\rightarrow,\leq,s_0)$ is ....

Let us remark that given a WSTS $\mathscr{S}=(S, \rightarrow, \leq,s_0)$, one may construct many \emph{guarded WSTSs} where every transition $s \rightarrow s'$ of $\mathscr{S}=(S, \rightarrow, \leq,s_0)$ is now guarded by a (monotone) formula $\phi$ defined by the grammar $\phi ::= s \geq u \mid s' \geq u \mid \phi \vee \phi \mid \phi \wedge \phi$ where $u \in S$. Since these formula are upward-compatible, the constructed transition system $\mathscr{S'}=(S, \rightarrow', \leq,s_0)$ is still a WSTS. 
%
For example, VASS allows such guards.
  %	\lor s' \geq x_2  	\lor...	\lor s' \geq x_p$ (facile dans les VASS et modèles habituels).

>>>>

% We prefer to use infix notation and $(s,a ,s')\in {\rightarrow} $ will be abbreviated as    $s  \xrightarrow{a}  s'$ to represent a transition from configuration $s$ to configuration $s'$ with label $a$. 
We prefer to use infix notation and write $s \rightarrow s'$ to denote a {\em transition} from state $s$ to state $s'$ (i.e., $ (s,s') \in  {\rightarrow} $).
We write $\rightarrow^{k}$, $\rightarrow^{+}$, $\rightarrow^{=}$, $\rightarrow^{*}$
for the $k$-step iteration of $\rightarrow$, its transitive closure, its reflexive closure, its reflexive and transitive closure.
Let $X,Y \subseteq S$ and $k \in \mathbb{N}$. We denote $X \longrightarrow^{*} Y$ (resp. $X \longrightarrow^{\leq k} Y$) if from all states $x \in X$ there exists a path (resp. of length smaller than $k$) that reaches a state $y \in Y$.
\noindent
The set of {\em (immediate) successors} of a state $s \in S$ is defined as 
 $\post(s) = \{ s' \in S \mid  ~ s \xrightarrow{} s'\}$. 
% $\post(s) = \{ s' \in S \mid  ~ \exists a \in \Sigma~ s \xrightarrow{a} s'\}$. 
A state without successors is called a {\em dead end}. 
The set of {\em (immediate) predecessors} of a state $s \in S$ is deﬁned as
 $\pred(s) = \{ s' \in S \mid  ~ s' \xrightarrow{} s\}$. 
% $\pred(s) = \{ s' \in S \mid  ~ \exists a \in \Sigma~ s' \xrightarrow{a} s\}$. 
By iterating $\pred$ and $\post$ we obtain  
$\post^n(s) = \{ s' \in S \mid  ~ s' \xrightarrow{}^n s\}$
and
$\pred^n(s) = \{ s' \in S \mid  ~ s' \xrightarrow{}^n s\}$.
However, we are generally more interested in
$\post^{\leq n}(s) = \bigcup_{1 \leq i \leq n} \post^i(s)$, $\post^*(s)= \bigcup_{1 \leq i} \post^i(s)$
and
$\pred^{\leq n}(s) = \bigcup_{1 \leq i \leq n} \pred^i(s)$ and $\pred^*(s) = \bigcup_{1 \leq i} \pred^i(s)$. \\

@@

\begin{definition}\cite{DBLP:journals/iandc/Finkel90,DBLP:journals/tcs/FinkelS01}
A {\em well-structured transition system} (WSTS)  $\mathscr{S}=(S, \rightarrow, \leq)$
is a transition system $(S, \rightarrow)$
equipped with a wqo ${\leq} \subseteq S \times S$ such that  
%  decidable ≤  wqo on S, i.e., for each two given states s, s ′ ∈ S, it is decidable whether s ≤ s ′ . 
the transition relation $ \rightarrow$ is (upward) compatible with $\leq$, i.e., for all 
$s_1, t_1 , s_2 \in S$
	with $s_1 \leq s_2$  and $s_1 \rightarrow t_1$, there exists 
	$t_2 \in S$ with 
	$t_1 \leq t_2$ and $s_2 \rightarrow^{*} t_2$.
\end{definition}

@@

We will use the following property: if $\mathscr{S}$ is a WSTS with strong compatibility and $U \subseteq S$ is upward-closed, then $\pred(U )$, $\pred^k(U )$ with $k\geq0$, and $\pred^*(U )$ are upward-closed \cite{DBLP:journals/tcs/FinkelS01}.
%\alain{étrange ces deux propositions, l'une suffirait}
\iffalse
Proof. Assume $s \in \pred (I )$. Then $s \rightarrow t$ for some $t \in I $. If now $s' \geq s$ then strong upward-compatibility entails that $s' \rightarrow t'$ for some $t' \geq t$. Then $t' \in I$ and $s' \in \pred(I )$.

\mathieu{On peut probablement enlever la Proposition $8$ alors. La proposition $9$ étant plus directement pertinente pour les lemmes/preuves (permet d'avoir itérativement $\pred^k(I)$ upward-closed if $I$ upward-closed - for all $k$).}
\fi
%
Now, we recall a simple condition that insures that a finite basis of $\pred^*(\uparrow s )$ is computable for every $s \in S$.

\begin{definition}\cite{DBLP:journals/tcs/FinkelS01,DBLP:journals/iandc/AbdullaCJT00}
A WSTS $\mathscr{S}=(S, \rightarrow, \leq)$ has {\em effective pred-basis} if there exists an algorithm accepting
any state $s \in S$ and returning $pb(s)$, a finite basis of $\uparrow \pred(\uparrow s)$.
\end{definition}

With the effective pred-basis hypothese, we obtain:

\begin{theorem}(ref)
A finite basis of $ \pred^*(U)$ is computable for any effective WSTS $\mathscr{S}=(S, \rightarrow, \leq)$ with effective pred-basis and any upward-closed set $U \subseteq S$ given with its finite basis $B_U$.
\end{theorem}

@@

% (ii) Given a finite set $A \subseteq S$ with $I =\uparrow A$, we can compute a finite basis $B$ of $I$.
% \alain{(ii) est bizarre car A est déjà une base finie de $\uparrow A$, et $\min(A)$ est la base minimale}
%			\end{fact}
For a WSTS $\mathscr{S}=(S, \rightarrow, \leq)$ and an upward-closed set $U  \subseteq S$, let us study the convergence of the sequence defined by $U_0=U$ and $U_k= U_{k-1} \cup \pred(U_{k-1})$ for $k \geq 1$. When $\mathscr{S}$ has the strong compatibility, the sets $U_k$ are upward-closed and $U_k \subseteq U_{k+1}$ so we know that the sequence $(U_k)_k$ converges. Let us define the \emph{index} of convergence of the sequence $U_k$ as the smallest $k_0$ s.t. $U_k = U_{k_0}$ for all $k \geq k_0$. We may compute $k_0$ and we then have:  $\pred^*(U) = U_{k_0}$. \\
%

\noindent
\subsection{Computing the maximal $X$-live subsystem}

Let $\mathscr{S}=(S, \rightarrow, \leq,s_0)$ be WSTS with initial state $s_0$ and $X \subseteq S$ be an upward-closed set given by its finite minimal basis $B_X$. We say that $\mathscr{S}=(S, \rightarrow, \leq,s_0)$ is $X$-live if from any reachable state $s \in \post^*(s_0)$, $s \xrightarrow{*} X$ is satisfied.  We may verify that  $\mathscr{S}=(S, \rightarrow, \leq,s_0)$ is $X$-live iff  $\mathscr{S}=(S, \rightarrow, \leq,s_0)$ is $\pred^*(X)$-live. The \emph{maximal $X$-live subsystem} of $(S,\rightarrow,\leq,s_0)$ is ....

Let us remark that given a WSTS $\mathscr{S}=(S, \rightarrow, \leq,s_0)$, one may construct many \emph{guard} systems where every transition $s \rightarrow s'$ of $\mathscr{S}$ is guarded by a (monotone) formula $\phi$ defined by the grammar $\phi ::= s \geq u \mid s' \geq u \mid \phi \vee \phi \mid \phi \wedge \phi$ where $u \in S$. Since guards restrict the original runs but don't create new ones, every guarded transition system $\mathscr{S'}$ is a subsystem of $\mathscr{S}$. Since these formula are upward-compatible, the guarded transition systems $\mathscr{S'}=(S, \rightarrow', \leq,s_0)$ are still WSTS.
%
For example, VASS allows such guards.
  %	\lor s' \geq x_2  	\lor...	\lor s' \geq x_p$ (facile dans les VASS et modèles habituels).
